---
title: "KT Protein Biodiversity Analysis"
author: "Artem Babaian"
date: "2026-03-30"
output:
  html_document:
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: show
    fig_width: 10
    fig_height: 6
editor_options: 
  chunk_output_type: inline
---

# Initialize Workspace

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = TRUE,
  fig.align = "center",
  cache = TRUE,
  cache.lazy = FALSE,
  results = "console"  # Default: send text output to console
)

# Load required packages
library(knitr)
library(dplyr)
library(ggplot2)
library(readr)
library(tibble)
library(scales)
library(kt)
library(data.table)

# Set theme for plots
theme_set(theme_minimal() +
          theme(plot.title = element_text(hjust = 0.5)))


```

# Package Building (Manual Run Only)

```{r package_build, eval=FALSE}
# Package development commands - run manually when needed
library(devtools)
library(roxygen2)

# Generate documentation from roxygen comments
document()

# Check package for issues
check()

# Install the package locally
# install()
roxygenize()

# Load the package after installation
library(kt)
```

# Fasta Files

## Load KT Sequence Data

```{r load_kt_data, results="inline"}
# Load KT1 sequences (clustered representatives)
kt1.df <- load_kt_fasta(
  file_path = "inst/extdata/kt1/kt1.fa",
  dataset_name = "kt1",
  package_data = FALSE
)

# Load KT0 sequences (pre-clustered sequences)
kt0.df <- load_kt_fasta(
  file_path = "inst/extdata/kt0/kt0.preclust.fa",
  dataset_name = "kt0",
  package_data = FALSE
)

```

## FASTA Data Visualization

```{r fasta_visualizations, fig.width=12, fig.height=8, results="console"}
# Family count comparison between KT1 and KT0
family_plot <- plot_family_count(
  KT1 = kt1.df,
  KT0 = kt0.df,
  title = "KT Family Distribution: KT1 vs KT0",
  top_n = 12
)
print(family_plot)
```

```{r organism_comparison, fig.width=12, fig.height=8, results="console"}
# Organism count comparison
organism_plot <- plot_organism_count(
  KT1 = kt1.df,
  KT0 = kt0.df,
  title = "Organism Distribution: KT1 vs KT0",
  top_n = 12,
  shorten_names = TRUE
)
print(organism_plot)
```

```{r length_distributions, fig.width=12, fig.height=6, results="console"}
# Sequence length distribution for all families
length_all_plot <- plot_family_length(
  data = kt1.df,
  family = "All",
  title = "KT1 Sequence Length Distribution - All Families",
  bins = 40,
  show_stats = TRUE
)
print(length_all_plot)
```

```{r length_by_family, fig.width=14, fig.height=10, results="console"}
# Length distribution separated by family
length_facet_plot <- plot_family_length(
  data = kt1.df,
  facet_families = TRUE,
  title = "KT1 Sequence Length Distribution by Family",
  bins = 25,
  show_stats = FALSE
)
print(length_facet_plot)
```

## Load UCLUST Clustering Data

```{r load_clustering_data, results="console"}
# Load sorted UCLUST clustering output (S records removed, sorted by cluster_id)
kt1.uclust <- load_uclust(
  file_path = "inst/extdata/kt1/kt1.id90.sorted.uc",
  include_seeds = FALSE,  # Using C records instead of S records for centroids
  include_hits = TRUE
)

```

## Clustering Analysis and Visualization

```{r cluster_size_analysis, fig.width=12, fig.height=8, results="console"}
# Analyze cluster size distribution
cluster_stats <- get_cluster_statistics(kt1.uclust)
#print("=== Clustering Summary ===")
#print(cluster_stats$summary)

# Create cluster size histogram
cluster_plot <- plot_cluster_counts(
  clustering_data = kt1.uclust,
  title = "KT1 Cluster Size Distribution",
  bins = 30,
  show_stats = TRUE
)
print(cluster_plot)
```

```{r cluster_analysis_detailed, fig.width=14, fig.height=10, results="console"}
# Detailed analysis: log scale and by family
cluster_log_plot <- plot_cluster_counts(
  clustering_data = kt1.uclust,
  title = "KT1 Cluster Sizes (Log Scale)",
  log_scale = TRUE,
  bins = 25,
  show_stats = FALSE
)

cluster_family_plot <- plot_cluster_counts(
  clustering_data = kt1.uclust,
  title = "Cluster Size Distribution",
  facet_by = "kt_family",
  bins = 20,
  show_stats = FALSE,
  min_cluster_size = 1
)

print(cluster_log_plot)
print(cluster_family_plot)
```

```{r cluster_identity_analysis, fig.width=12, fig.height=8, results="console"}
# Cluster size vs identity relationship
identity_plot <- plot_cluster_size_identity(
  clustering_data = kt1.uclust,
  min_cluster_size = 2,
  color_by = "kt_family"
)
print(identity_plot)

# Display family-wise clustering statistics
print("=== Family-wise Clustering Statistics ===")
print(cluster_stats$family_stats)

print("\n=== Top 10 Largest Clusters ===")
print(cluster_stats$top_clusters)
```

## Load All-vs-All Alignment Data

```{r load_alignment_data, results="console"}
# Load USEARCH allpairs alignment output
kt1.alignments <- load_kt_alignment(
  file_path = "inst/extdata/kt1/kt1.aln",
  include_self = FALSE,  # Remove self-alignments
  min_identity = 25,     # Minimum 25% identity (as per USEARCH parameters)
  min_length = 50        # Minimum alignment length filter
)

print("All-vs-All Alignment Data Loaded:")
#tr(kt1.alignments)

# Get alignment summary statistics
alignment_summary <- summarize_alignment(kt1.alignments)
#print("Alignment Summary:")
#print(alignment_summary)

# Check for reciprocal alignments
#print("\n=== Reciprocal Alignment Analysis ===")
reciprocal_pairs <- kt1.alignments[, .(
  pair1 = paste(pmin(query_accession, subject_accession),
                pmax(query_accession, subject_accession), sep = "_"),
  query = query_accession,
  subject = subject_accession,
  identity = percent_identity
)]

# Count unique pairs vs total alignments
unique_pairs <- uniqueN(reciprocal_pairs$pair1)
total_alignments <- nrow(reciprocal_pairs)

#print(paste("Total alignments:", total_alignments))
#print(paste("Unique protein pairs:", unique_pairs))
#print(paste("Reciprocal ratio:", round(total_alignments / unique_pairs, 2)))

if (total_alignments / unique_pairs > 1.5) {
  print("Most pairs have reciprocal alignments - graph should be undirected")
} else {
  print("Few reciprocal alignments found - data appears to be already deduplicated")
}

```

## High-Quality Alignments Analysis

```{r high_quality_alignments, results="console"}
# Filter for high-quality alignments
hq_alignments <- filter_alignments(
  kt1.alignments,
  min_identity = 30,
  min_length = 100,
  max_evalue = 1e-10
)

print(paste("High-quality alignments:", nrow(hq_alignments)))

# Summarize by family pairs
family_stats <- summarize_alignment(hq_alignments, by_family = TRUE)
# print("Top family pair alignments:")
print(family_stats)

```

# Network Analysis and Visualization

## Setup Network Packages

```{r setup_network_packages, results="console"}
# Load packages
library(RCy3)
library(igraph)
library(viridis)

# Check if Cytoscape is running
cytoscapePing()
```

## Create Network Graph from Alignments

```{r create_network_graph, results="console"}
# Create network graph from high-quality alignments
g <- create_kt_graph(
  alignment_data = hq_alignments,
  layout_method = "kamada.kawai",
  layout_iterations = 10000
)
```

## Add KT0 Cluster Member Counts

```{r add_kt0_member_counts, results="console"}

# Count KT0 sequences per centroid from clustering data
# Need to count both: 1) centroid itself if from kt0, 2) H records from kt0

# Get centroids and check if they're from kt0
centroids <- kt1.uclust[record_type == "C", .(centroid_accession = accession,
                                              cluster_id,
                                              centroid_from_kt0 = (search_group == "kt"))]

# Get members (H records) from kt0
members <- kt1.uclust[record_type == "H", .(member_accession = accession,
                                           centroid_accession = sub(".*?([^;]+)$", "\\1", target_label),
                                           cluster_id, search_group)]

# Count KT0 members per centroid (H records only)
kt0_hit_counts <- members[search_group == "kt", .N, by = centroid_accession]
setnames(kt0_hit_counts, c("centroid_accession", "kt0_hit_count"))

# Merge centroids with hit counts
kt0_counts_complete <- merge(centroids[, .(centroid_accession, centroid_from_kt0)],
                            kt0_hit_counts, by = "centroid_accession", all.x = TRUE)
kt0_counts_complete[is.na(kt0_hit_count), kt0_hit_count := 0]

# Total kt0 count = centroid (if from kt0) + hit members from kt0
kt0_counts_complete[, kt0_member_count := as.integer(centroid_from_kt0) + kt0_hit_count]

print("KT0 member counts per centroid (including centroid itself):")
print(kt0_counts_complete[order(-kt0_member_count), .(centroid_accession, centroid_from_kt0, kt0_hit_count, kt0_member_count)])

# Simplify to just the columns we need
kt0_counts_complete <- kt0_counts_complete[, .(centroid_accession, kt0_member_count)]

print(paste("Total centroids:", nrow(kt0_counts_complete)))
print(paste("Centroids with KT0 members:", sum(kt0_counts_complete$kt0_member_count > 0)))
print(paste("Total KT0 sequences clustered:", sum(kt0_counts_complete$kt0_member_count)))
```

## Apply Visual Properties

```{r apply_visual_properties, results="console"}
# Apply visual styling to the graph
g <- apply_kt_graph_visuals(
  graph = g,
  edge_color_by = "identity",
  color_palette = viridis,
  edge_width_range = c(0.5, 3),
  node_size = 8,
  node_color_by = "family"  # Color nodes by KT family
)

# Add KT0 member count as a node attribute
library(igraph)
node_names <- V(g)$name
kt0_counts_for_nodes <- kt0_counts_complete$kt0_member_count[match(node_names, kt0_counts_complete$centroid_accession)]
kt0_counts_for_nodes[is.na(kt0_counts_for_nodes)] <- 0  # Nodes not in clustering data get count of 0

V(g)$kt0_member_count <- kt0_counts_for_nodes

print("Visual properties applied to network")
print(paste("Added KT0 member counts to", length(V(g)$kt0_member_count), "nodes"))
print(paste("Nodes with KT0 members:", sum(V(g)$kt0_member_count > 0)))

# Display summary of KT0 member count distribution
kt0_count_summary <- summary(V(g)$kt0_member_count)
print("KT0 member count distribution across network nodes:")
print(kt0_count_summary)
```

## Network Statistics

```{r network_statistics, results="console"}
# Get comprehensive network statistics
stats <- get_kt_graph_stats(g)

# Update graph with community information
g <- stats$graph_updated

# Display results
print("=== Network Statistics ===")
print(paste("Nodes:", stats$basic_stats$nodes))
print(paste("Edges:", stats$basic_stats$edges))
print(paste("Density:", stats$basic_stats$density))
print(paste("Clustering coefficient:", stats$basic_stats$clustering))
print(paste("Average path length:", stats$basic_stats$avg_path_length))
  
print("\n=== Degree Distribution ===")
print(paste("Average degree:", stats$degree_stats$avg_degree))
print(paste("Maximum degree:", stats$degree_stats$max_degree))
print("Top 10 most connected accessions:")
print(stats$degree_stats$top_nodes)

print("\n=== Community Detection ===")
print(paste("Communities detected:", stats$community_stats$num_communities))
print(paste("Modularity:", stats$community_stats$modularity))

print("\n=== Family-based Analysis ===")
print("Top families by node count:")
print(head(stats$family_stats$family_counts, 10))

print(paste("Intra-family edges:", stats$family_stats$intra_family_edges))
print(paste("Inter-family edges:", stats$family_stats$inter_family_edges))
print(paste("Inter/Intra ratio:", stats$family_stats$inter_intra_ratio))

print("\nTop families by average node degree:")
print(stats$family_stats$top_families_by_degree)

print("\n=== KT0 Cluster Member Analysis ===")
kt0_member_stats <- list(
  total_nodes_with_kt0 = sum(V(g)$kt0_member_count > 0),
  total_kt0_members = sum(V(g)$kt0_member_count),
  avg_kt0_per_node = round(mean(V(g)$kt0_member_count), 2),
  max_kt0_per_node = max(V(g)$kt0_member_count),
  nodes_with_most_kt0 = names(V(g))[order(V(g)$kt0_member_count, decreasing = TRUE)[1:5]]
)

print(paste("Nodes with KT0 members:", kt0_member_stats$total_nodes_with_kt0))
print(paste("Total KT0 sequences represented:", kt0_member_stats$total_kt0_members))
print(paste("Average KT0 members per node:", kt0_member_stats$avg_kt0_per_node))
print(paste("Maximum KT0 members in single node:", kt0_member_stats$max_kt0_per_node))
print("Top 5 nodes by KT0 member count:")
for(i in 1:5) {
  node_name <- kt0_member_stats$nodes_with_most_kt0[i]
  count <- V(g)$kt0_member_count[V(g)$name == node_name]
  family <- V(g)$family[V(g)$name == node_name]
  print(paste(" ", i, ". ", node_name, " (", family, "): ", count, " KT0 members", sep = ""))
}
```

## Network Visualization

```{r network_plots, results="console"}
# Create summary plot
plot_kt_graph_summary(g, stats)
```

```{r network_main_plot, fig.width=12, fig.height=10, results="console"}
# Main network plot
plot_kt_graph(g,
             main = "KT Protein Alignment Network",
             show_labels = FALSE,
             legend_position = "topright")
```

```{r network_kt0_plot, fig.width=12, fig.height=10, results="console"}
# Network plot with nodes sized by KT0 member count
# Create a copy of the graph for this visualization
g_kt0 <- g

# Scale node sizes based on KT0 member count (minimum size 2, maximum size 15)
kt0_counts <- V(g_kt0)$kt0_member_count
if (max(kt0_counts) > 0) {
  node_sizes <- 2 + (kt0_counts / max(kt0_counts)) * 13
} else {
  node_sizes <- rep(5, length(kt0_counts))
}
V(g_kt0)$size <- node_sizes

# Create color palette for KT0 counts
if (max(kt0_counts) > 0) {
  # Use a color scale from light to dark blue based on KT0 count
  color_breaks <- seq(0, max(kt0_counts), length.out = 10)
  node_colors <- colorRampPalette(c("lightgray", "darkblue"))(10)[cut(kt0_counts, breaks = color_breaks, include.lowest = TRUE)]
  V(g_kt0)$color <- node_colors
} else {
  V(g_kt0)$color <- "lightgray"
}

# Plot the network
plot(g_kt0,
     main = "KT Network: Node Size by KT0 Member Count",
     vertex.label = ifelse(V(g_kt0)$kt0_member_count > quantile(V(g_kt0)$kt0_member_count, 0.95), V(g_kt0)$name, NA),
     vertex.label.cex = 0.6,
     vertex.label.color = "black",
     edge.color = "lightgray",
     edge.width = 0.5,
     layout = cbind(V(g_kt0)$x, V(g_kt0)$y))

# Add legend
if (max(kt0_counts) > 0) {
  legend("topright",
         legend = c(paste("Max:", max(kt0_counts), "KT0"),
                   paste("Min:", min(kt0_counts), "KT0")),
         col = c("darkblue", "lightgray"),
         pch = 19,
         pt.cex = c(2, 0.5),
         title = "KT0 Members")
}
```

## Export to Cytoscape

```{r export_to_cytoscape, eval=FALSE, results="console"}
# TODO: Calculate a MST to determine the Layout of the graph and save
#       those X,Y coordinates as the basis of the layout
# Export to Cytoscape with fallback to CSV files
export_result <- export_kt_to_cytoscape(
  graph = g,
  network_name = "KT_Alignment_Network",
  collection_name = "KT Analysis",
  style_name = "KT_Network_Style",
  edge_color_attribute = "identity"
)

print(paste("Export method:", export_result$method))
if (!is.null(export_result$files)) {
  print("Files created:")
  print(export_result$files)
}

```

## Import M1 Pro File
```{r import_m1_pro, results="console"}
#pro.df <- read_pro('inst/extdata/kt1.M1_Logan.pro')

#View(pro.df[ pro.df$sseqid == 'kt1.kt;M1;Saccharomyces_cerevisiae;NP_044402.1', ])


```


## TODO

### In Progress đź”„
- [ ] Install and load Biostrings package for sequence analysis
- [ ] Add error handling for missing data files
- [ ] Create alignment quality distribution plots

### Future Enhancements đź”®
- [ ] Add statistical tests for biodiversity comparisons
- [ ] Implement phylogenetic analysis pipeline
- [ ] Analyze reciprocal best hits from alignment data
- [ ] Create interactive network visualizations with plotly/visNetwork
- [ ] Add sequence similarity clustering validation
- [ ] Implement protein domain analysis integration
- [ ] Add comparative genomics analysis functions

## Node Degree Analysis for MLus Centroids

```{r mlus_degree_histogram, fig.width=10, fig.height=6, results="console"}
# Extract node information from the network graph
node_data <- data.frame(
  accession = V(g)$name,
  family = V(g)$family,
  degree = degree(g),
  kt0_member_count = V(g)$kt0_member_count,
  stringsAsFactors = FALSE
)

# Check available families
print("Available KT families in network:")
print(sort(table(node_data$family), decreasing = TRUE))

# Use all families present in the data
all_families <- unique(node_data$family)
all_families <- all_families[!is.na(all_families) & all_families != ""]

# Filter out families with very few nodes (less than 3) for cleaner visualization
family_counts_all <- table(node_data$family)
families_with_enough_nodes <- names(family_counts_all[family_counts_all >= 3])

# Use families with sufficient representation
family_nodes <- node_data[node_data$family %in% families_with_enough_nodes, ]

print(paste("Total families in data:", length(all_families)))
print(paste("Families with â‰Ą3 nodes (shown in plot):", length(families_with_enough_nodes)))

# Print counts per family (sorted by count)
family_counts <- table(family_nodes$family)
family_counts_sorted <- sort(family_counts, decreasing = TRUE)
print("Node counts by family (top 15):")
print(head(family_counts_sorted, 15))

# Create faceted histogram comparing degree distributions across ALL families
if (nrow(family_nodes) > 0) {
  # Calculate mean degree for each family for vertical lines
  family_means <- family_nodes %>%
    group_by(family) %>%
    summarise(mean_degree = mean(degree), .groups = 'drop')

  # Determine optimal number of columns based on number of families
  num_families <- length(families_with_enough_nodes)
  ncol_facets <- if (num_families <= 6) 2 else if (num_families <= 12) 3 else 4

  family_histogram <- ggplot(family_nodes, aes(x = degree)) +
    geom_histogram(bins = 12,
                   fill = "steelblue",
                   alpha = 0.7,
                   color = "black") +
    geom_vline(data = family_means,
               aes(xintercept = mean_degree),
               color = "red",
               linetype = "dashed",
               size = 0.6) +
    facet_wrap(~ family, scales = "free_y", ncol = ncol_facets) +
    labs(
      title = "Node Degree Distribution Across All KT Families",
      x = "Node Degree",
      y = "Count of Centroid Nodes",
      subtitle = paste("Showing", length(families_with_enough_nodes),
                      "families with â‰Ą3 nodes each")
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 8),
      strip.text = element_text(size = 9, face = "bold"),
      panel.spacing = unit(0.5, "lines")
    )

  print(family_histogram)

  # Print detailed statistics for all families (sorted by mean degree)
  print("\n=== All Family Degree Statistics (sorted by mean degree) ===")
  family_stats <- family_nodes %>%
    group_by(family) %>%
    summarise(
      count = n(),
      min_degree = min(degree),
      max_degree = max(degree),
      mean_degree = round(mean(degree), 2),
      median_degree = median(degree),
      sd_degree = round(sd(degree), 2),
      .groups = 'drop'
    ) %>%
    arrange(desc(mean_degree))

  print(family_stats)

  # Show top degree node for each family (condensed view)
  print("\n=== Highest degree node per family ===")
  top_per_family <- family_nodes %>%
    group_by(family) %>%
    arrange(desc(degree)) %>%
    slice_head(n = 1) %>%
    select(family, accession, degree, kt0_member_count) %>%
    arrange(desc(degree))

  print(top_per_family)

  # Create a second plot focusing on the most populous families
  top_families <- names(head(family_counts_sorted, 8))
  if (length(top_families) > 0) {
    top_family_nodes <- family_nodes[family_nodes$family %in% top_families, ]

    print(paste("\n=== Creating detailed view for top", length(top_families), "most populous families ==="))

    top_family_means <- top_family_nodes %>%
      group_by(family) %>%
      summarise(mean_degree = mean(degree), .groups = 'drop')

    top_families_plot <- ggplot(top_family_nodes, aes(x = degree)) +
      geom_histogram(bins = 15,
                     fill = "darkgreen",
                     alpha = 0.7,
                     color = "black") +
      geom_vline(data = top_family_means,
                 aes(xintercept = mean_degree),
                 color = "red",
                 linetype = "dashed",
                 size = 0.8) +
      facet_wrap(~ family, scales = "free_y", ncol = 2) +
      labs(
        title = "Detailed View: Top 8 Most Populous KT Families",
        x = "Node Degree",
        y = "Count of Centroid Nodes",
        subtitle = "Higher resolution view of families with most representatives"
      ) +
      theme_minimal() +
      theme(
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 11, face = "bold")
      )

    print(top_families_plot)
  }

} else {
  print("No family nodes found with sufficient representation.")
  print("All available families:")
  print(sort(unique(node_data$family)))
}
```

## M28 Family Investigation

```{r m28_investigation, results="console"}
print("=== Investigating M28 Family Representation ===")

# Check 1: Is M28 in the original KT1 FASTA data?
print("\n1. Checking KT1 FASTA data for M28/M28-like families:")
kt1_families <- table(kt1.df$kt_family)
m28_related <- kt1_families[grepl("M28|m28", names(kt1_families), ignore.case = TRUE)]
if (length(m28_related) > 0) {
  print("Found M28-related families in KT1:")
  print(m28_related)
} else {
  print("No M28 family found in KT1 FASTA data")
  print("All families in KT1 (sorted by count):")
  print(sort(kt1_families, decreasing = TRUE))
}

# Check 2: Is M28 in the KT0 data?
print("\n2. Checking KT0 FASTA data for M28:")
kt0_families <- table(kt0.df$kt_family)
m28_related_kt0 <- kt0_families[grepl("M28|m28", names(kt0_families), ignore.case = TRUE)]
if (length(m28_related_kt0) > 0) {
  print("Found M28-related families in KT0:")
  print(m28_related_kt0)
} else {
  print("No M28 family found in KT0 FASTA data")
}

# Check 3: Look for similar names that might be M28
print("\n3. Looking for families with 'M' and numbers that might be M28:")
all_families_combined <- c(names(kt1_families), names(kt0_families))
m_families <- unique(all_families_combined[grepl("M[0-9]", all_families_combined)])
print("Families starting with M followed by numbers:")
print(sort(m_families))

# Check 4: Are there any M28 sequences in clustering data?
print("\n4. Checking clustering data for M28:")
if (exists("kt1.uclust")) {
  # Extract family information from clustering data if available
  clustering_accessions <- unique(c(kt1.uclust$accession, kt1.uclust$target_label))

  # Check if any M28 accessions in clustering
  m28_in_clustering <- any(grepl("M28|m28", clustering_accessions, ignore.case = TRUE))
  print(paste("M28 sequences in clustering data:", m28_in_clustering))

  if (m28_in_clustering) {
    m28_cluster_seqs <- clustering_accessions[grepl("M28|m28", clustering_accessions, ignore.case = TRUE)]
    print("M28 sequences found:")
    print(head(m28_cluster_seqs))
  }
}

# Check 5: Are there M28 sequences in alignment data?
print("\n5. Checking alignment data for M28:")
if (exists("kt1.alignments")) {
  alignment_accessions <- unique(c(kt1.alignments$query_accession, kt1.alignments$subject_accession))
  m28_in_alignments <- any(grepl("M28|m28", alignment_accessions, ignore.case = TRUE))
  print(paste("M28 sequences in alignment data:", m28_in_alignments))

  if (m28_in_alignments) {
    m28_align_seqs <- alignment_accessions[grepl("M28|m28", alignment_accessions, ignore.case = TRUE)]
    print("M28 sequences in alignments:")
    print(head(m28_align_seqs))
  }
}

# Check 6: Are there M28 nodes in the network graph?
print("\n6. Checking network graph for M28:")
if (exists("g")) {
  graph_accessions <- V(g)$name
  graph_families <- V(g)$family

  m28_in_graph <- any(grepl("M28|m28", graph_accessions, ignore.case = TRUE)) ||
                  any(grepl("M28|m28", graph_families, ignore.case = TRUE))
  print(paste("M28 sequences in network graph:", m28_in_graph))

  if (m28_in_graph) {
    m28_graph_nodes <- graph_accessions[grepl("M28|m28", graph_accessions, ignore.case = TRUE)]
    m28_family_nodes <- which(grepl("M28|m28", graph_families, ignore.case = TRUE))
    print("M28 nodes in graph:")
    if (length(m28_graph_nodes) > 0) print(head(m28_graph_nodes))
    if (length(m28_family_nodes) > 0) print(paste("Family nodes:", length(m28_family_nodes)))
  }
}

# Check 7: Look for case variations or alternative naming
print("\n7. Looking for alternative M28 naming patterns:")
all_families_unique <- unique(c(names(kt1_families), names(kt0_families)))
potential_m28 <- all_families_unique[grepl("28|M.*28|m.*28", all_families_unique, ignore.case = TRUE)]
if (length(potential_m28) > 0) {
  print("Families containing '28':")
  print(potential_m28)
} else {
  print("No families containing '28' found")
}

# Check 8: Display family parsing from sample sequences
print("\n8. Sample family parsing to check format:")
if (nrow(kt1.df) > 0) {
  print("Sample KT1 headers and parsed families:")
  sample_kt1 <- head(kt1.df[, c("kt_family", "kt_family")], 10)
  print(sample_kt1)
}

print("\n=== Summary ===")
print("If M28 is not found in any of the above checks, it likely means:")
print("1. M28 sequences are not present in the source FASTA files")
print("2. M28 sequences failed quality filters during processing")
print("3. M28 is named differently than expected")
print("4. M28 sequences didn't form alignments meeting the criteria")
print("5. M28 sequences are present but didn't become cluster centroids")
```
```



































